Stroboscopic Doppler tomography of FO Aqr

### 2013-12-08 paper/bptectono.tex
figures much better

\documentclass[draft, 12pt]{article}
\usepackage[english]{babel}
\usepackage[final]{graphicx}
\usepackage{multirow}
\usepackage{color}
\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.png}

@@

\title{Seismic parameter estimation and the Canadian crust}
\author{Ben Postlethwaite}

%% ------------------------------------------------------------------------ %%
\begin{document}

\begin{abstract}

@@

%% -----------------------------------------------%%
\section{Introduction}
\section{Introduction}

@@

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{reflectedPhases}
  \caption{Schematic diagram illustrating geometry of phases for the velocity contrast representing the Moho}
  \label{fig:reflectedPhases}
\end{figure}

@@

\begin{table}
  \begin{tabular}{ l l l l l l l}
    \cline{4-7}
    & & \multicolumn{2}{ c }{Correlation Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
    \hline
    Study Authors             & Region         & Num. of Stns &  $H$ & $R$  &$H$(km)& $R$ \\
    \hline
    Thompson et. al. (2010)   & Northern Canadian Sheild & 29 & 0.97 & 0.70 & 0.78 & 0.017 \\
    Darbyshire et. al. (2007) & Superior Province        & 10 & 0.95 & 0.43 & 1.71 & 0.037 \\
    Eaton et. al. (2006)      & Grenville Orogen         & 26 & 0.90 & 0.61 & 1.03 & 0.033 \\
    \hline
  \end{tabular}
  \caption{Comparison of $R$ and $H$ estimates with three published studies}
\label{table:comparison}

@@

\noindent As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended along the kinematic curve in the $H$ - $V_P$ space (Figure \ref{fig:HVp_ULM}). This trade-off contributes to the low correlation in $H$ between the two methods. Higher correlation in this quantity can be restored by dividing out $V_P$ from $H$. The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in an other to maintain travel-time equality, which has the result of a well defined misfit surface (Figures \ref{fig:HR_ULM}-\ref{fig:RVp_ULM}.

For all but the cleanest stations, the trade-off between $H$ and $V_P$ renders recovery of $V_P$ unfeasible using the full grid approach. A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting less resolution in the reflected phases, illustrates this difficulty. RFs are arranged vertically, ordered by increasing slowness, and shown in figures \ref{fig:RF_ULM} and \ref{fig:RF_DORN}. Cross sections of $s(H,R,V_P)$ through each coordinate plane are displayed for ULM and DORN in figures \ref{fig:HR_ULM}-\ref{fig:HVp_ULM}) and figures \ref{fig:HR_DORN}-\ref{fig:HVp_DORN}. The lower resolution apparent in the $H$ - $V_P$ cross section for DORN translates into a bootstrap error of $V_P \pm 0.64 {\rm km/s}$. Station ULM, with cleaner data, has less trade-off between $H$ and $V_P$ as indicated by the corresponding cross section and the lower bootstrap error of $V_P \pm 0.16 {\rm km/s}$.

Active source experiments located near broadband seismic stations provide independent crustal thickness and {\it P} wave velocity estimates to test against data derived from the full parameter search method. We select stations with low error in $V_P$ and active source experiment locations within a 100 km radius. If several active source experiments are located within this radius, estimates are averaged. Four stations (ARVN, TYNO, ULM and WHY) meet these criteria and the parameter estimates are presented in figure \ref{fig:activeVp}-\ref{fig:activeH}. Estimates computed from the full grid approach are roughly 0.25 km/s higher than the active source data. This difference is consistent across all four stations and is exemplified by a 99\% correlation, suggesting possible bias. However, the limited number of data means this bias is poorly constrained.

@@

[[ SOURCES REGARDING BIAS ]] [[ WHAT DOES THIS MEAN, IS VP USABLE FOR THE HIGHEST QUALITY STATIONS?]]

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{RF_ULM}
  \caption{}
  \label{fig:RF_ULM}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{stackHR_ULM}
  \caption{}
  \label{fig:HR_ULM}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{stackRVp_ULM}
  \caption{}
  \label{fig:RVp_ULM}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{stackHVp_ULM}
  \caption{}
  \label{fig:HVp_ULM}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{RF_DORN}
  \caption{}
  \label{fig:RF_DORN}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{stackHR_DORN}
  \caption{}
  \label{fig:HR_DORN}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{stackRVp_DORN}
  \caption{}
  \label{fig:RVp_DORN}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{stackHVp_DORN}
  \caption{}
  \label{fig:HVp_DORN}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{activeSourceCompVp}
  \caption{\it{P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
  \label{fig:activeVp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{activeSourceCompH}
  \caption{Crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
  \label{fig:activeH}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{conradPCA}
  \caption{The SVD singular vectors (principal components) of the station depth profile matrix. Each principal component represents a mode that captures some part of the total variance in the dataset. Vertical lines are the axis each profile and vector is normalized about. A grey horizontal window highlights a region of low energy.}
  \label{fig:conradPCA}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{discontinuity}
  \caption{}
  \label{fig:disc}
\end{figure}

@@

Table \ref{table:regionParameters} displays the results of our regionalization in $R$ and $H$. The weighted average crustal thickness for all stations (Canada) is 37 km, about two kilometres thinner than the average for the Canadian Shield. Crust 1.0 provides a slightly thicker crustal average for Canada (38.1 km) and a slightly thinner average for the Shield (38.3 km). The Slave Province averages are similar to those for the Churchill province with a crustal thickness and seismic velocity ratio comparable to Crust 1.0. Moving south and east from the Churchill Province through the Superior Province into the Grenville Province, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$, figures \ref{map:Hmap}-\ref{map:Rmap}. The weighted $R$ average is 1.73 in the Churchill Province increasing slightly to 1.74 in the Superior and jumping to 1.78 in the Grenville. In a similar fashion, the weighted average of $H$ increases from 38.9 km in the Churchill Province  to 39.7 in the Superior Province and 41.7 km in the Grenville Province. The southeastern trend towards increasing values is also expressed in weighted averaged active source $V_P$ data as well as as the Crust 1.0 $V_P$ data.  The Crust 1.0 model is characterized by slightly thinner crust though the southward thickening trend is noticeable. The trend in the seismic velocity ratio is not, however, as apparent.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{stationMap}
  \caption{}
  \label{map:stationMap}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{Hmap}
  \caption{}
  \label{map:Hmap}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{Rmap}
  \caption{}
  \label{map:Rmap}
\end{figure}

@@

\section{Discussion}

@@

\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{discontinuityInterp}
 \caption{Left: Vertical station depth profiles interpolated to a common Moho. Right: Normalized stack of depth profiles, grey line shows mean}
 \label{fig:discInterp}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{lithology}
  \caption{$V_P$ and $V_S$ estimates for common crust lithologies and for major Canadian-Shield provinces}
  \label{fig:lith}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{composition}
  \caption{Comparison of major crustal elements from different studies. All estimates have been normalized to the results computed in this study.}
  \label{fig:composition}
\end{figure}

>>>>

\documentclass{elsart}
\usepackage{graphicx,multirow,color,natbib,amssymb}
\journal{Tectonophysics}

\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.png}

@@

\begin{document}
\begin{frontmatter}

\title{Stroboscopic Doppler tomography of FO Aqr}
%% \author[South]{T.R. Marsh\corauthref{cor}}
%% \corauth[cor]{Corresponding author.}
%% \ead{trm@astro.soton.ac.uk}
%% \author[Oxford]{S.R. Duck\thanksref{now}}
%% \thanks[now]{Present address: Systems Engineering and Assessment Ltd., Beckington Castle, PO Box 800, Bath BA3 6TB, UK.}
%% \ead{srd@sea.co.uk}
%% \address[South]{University of Southampton, Department of Physics, Highfield, Southampton SO17 1BJ, UK}
%% \address[Oxford]{University of Oxford, Department of Physics, Nuclear Physics Laboratory, Keble Road, Oxford, OX1 3RH, UK}

%% ------------------------------------------------------------------------ %%
\begin{abstract}

@@

\end{frontmatter}
%% -----------------------------------------------%%
\section{Introduction}

@@

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{reflectedPhases.pdf}
  \caption{Schematic diagram illustrating geometry of phases for the velocity contrast representing the Moho}
  \label{fig:reflectedPhases}
\end{figure}

@@

\begin{table}
  \begin{tabular}{ l l l l l l l}
    \cline{4-7}
    & & & \multicolumn{2}{ c }{Correlation Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
    \hline
    Study Authors             & Region         & Num. of Stns &  $H$ & $R$  &$H$(km)& $R$ \\
    \hline
    Thompson et. al. (2010)   & Northern Canadian Sheild & 29 & 0.97 & 0.70 & 0.78 & 0.017 \\
    Darbyshire et. al. (2007) & Superior Province        & 10 & 0.95 & 0.43 & 1.71 & 0.037 \\
    Eaton et. al. (2006)      & Grenville Orogen         & 26 & 0.90 & 0.61 & 1.03 & 0.033 \\
    \hline
  \end{tabular}
  \caption{Comparison of $R$ and $H$ estimates with three published studies}
\label{table:comparison}

@@

\noindent As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended along the kinematic curve in the $H$ - $V_P$ space (Figure \ref{fig:kcurves} a). This trade-off contributes to the low correlation in $H$ between the two methods. Higher correlation in this quantity can be restored by dividing out $V_P$ from $H$. The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in an other to maintain travel-time equality, which has the result of a well defined misfit surface (Figure \ref{fig:kcurves} b-f).

For all but the cleanest stations, the trade-off between $H$ and $V_P$ renders recovery of $V_P$ unfeasible using the full grid approach. A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting less resolution in the reflected phases, illustrates this difficulty. RFs are arranged vertically, ordered by increasing slowness, and shown in figure \ref{fig:rfs}. Cross sections of $s(H,R,V_P)$ through each coordinate plane are displayed for ULM and DORN in figure \ref{fig:kcurves}. The lower resolution apparent in the $H$ - $V_P$ cross section for DORN translates into a bootstrap error of $V_P \pm 0.64 {\rm km/s}$. Station ULM, with cleaner data, has less trade-off between $H$ and $V_P$ as indicated by the corresponding cross section and the lower bootstrap error of $V_P \pm 0.16 {\rm km/s}$.

Active source experiments located near broadband seismic stations provide independent crustal thickness and {\it P} wave velocity estimates to test against data derived from the full parameter search method. We select stations with low error in $V_P$ and active source experiment locations within a 100 km radius. If several active source experiments are located within this radius, estimates are averaged. Four stations (ARVN, TYNO, ULM and WHY) meet these criteria and the parameter estimates are presented in figure \ref{fig:vpactive}. Estimates computed from the full grid approach are roughly 0.25 km/s higher than the active source data. This difference is consistent across all four stations and is exemplified by a 99\% correlation, suggesting possible bias. However, the limited number of data means this bias is poorly constrained.

@@

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{kinematicCurves.eps}
  \caption{}
  \label{fig:kcurves}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{rfDORNULM.eps}
  \caption{}
  \label{fig:rfs}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{vpactive.eps}
  \caption{Top: {\it P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 1.0. Bottom: Crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
  \label{fig:vpactive}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{conradPCA.eps}
  \caption{The SVD singular vectors (principal components) of the station depth profile matrix. Each principal component represents a mode that captures some part of the total variance in the dataset. Vertical lines are the axis each profile and vector is normalized about. A grey horizontal window highlights a region of low energy.}
  \label{fig:conradPCA}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{discontinuity.eps}
  \caption{}
  \label{fig:disc}
\end{figure}

@@

Table \ref{table:regionParameters} displays the results of our regionalization in $R$ and $H$. The weighted average crustal thickness for all stations (Canada) is 37 km, about two kilometres thinner than the average for the Canadian Shield. Crust 1.0 provides a slightly thicker crustal average for Canada (38.1 km) and a slightly thinner average for the Shield (38.3 km). The Slave Province averages are similar to those for the Churchill province with a crustal thickness and seismic velocity ratio comparable to Crust 1.0. Moving south and east from the Churchill Province through the Superior Province into the Grenville Province, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$. The weighted $R$ average is 1.73 in the Churchill Province increasing slightly to 1.74 in the Superior and jumping to 1.78 in the Grenville. In a similar fashion, the weighted average of $H$ increases from 38.9 km in the Churchill Province  to 39.7 in the Superior Province and 41.7 km in the Grenville Province. The southeastern trend towards increasing values is also expressed in weighted averaged active source $V_P$ data as well as as the Crust 1.0 $V_P$ data.  The Crust 1.0 model is characterized by slightly thinner crust though the southward thickening trend is noticeable. The trend in the seismic velocity ratio is not, however, as apparent.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{stationMap.pdf}
  \caption{}
  \label{map:stationMap}
\end{figure}

@@

@@

\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{discontinuityInterp.eps}
 \caption{Left: Vertical station depth profiles interpolated to a common Moho. Right: Normalized stack of depth profiles, grey line shows mean}
 \label{fig:discInterp}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{lithology.eps}
  \caption{$V_P$ and $V_S$ estimates for common crust lithologies and for major Canadian-Shield provinces}
  \label{fig:lith}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{composition.eps}
  \caption{Comparison of major crustal elements from different studies. All estimates have been normalized to the results computed in this study.}
  \label{fig:composition}
\end{figure}
